---
title: "Frequency analysis of a French Motor Third Party Liability dataset"
date: '`r Sys.Date()`'
date-format: "MMMM, YYYY"
author: "Julien Siharath"
categories: [Poisson, Public liability, Claim frequency, France, freMTPL]
vignette: >
  %\VignetteIndexEntry{Poisson frequency analysis}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}     
---


# Introduction {#sec-poisson-introduction}

```{r setup, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Session Settings"
#| label: load-packages

# Graphs----
face_text='plain'
face_title='plain'
size_title = 14
size_text = 11
legend_size = 11

global_theme <- function() {
  theme_minimal() %+replace%
    theme(
      text = element_text(size = size_text, face = face_text),
      legend.position = "bottom",
      legend.direction = "horizontal", 
      legend.box = "vertical",
      legend.key = element_blank(),
      legend.text = element_text(size = legend_size),
      axis.text = element_text(size = size_text, face = face_text), 
      plot.title = element_text(
        size = size_title, 
        hjust = 0.5
      ),
      plot.subtitle = element_text(hjust = 0.5)
    )
}

# Outputs
options("digits" = 2)
```

```{css setup, echo = FALSE}
.justify {
  text-align: justify !important
}
```

::: callout-tip
### In Brief

The objective of this vignette is to demonstrate the application of
Poisson regression in analyzing insurance data, specifically focusing on
the `freMTPLfreq` and `freMTPLsev` datasets from @charpentierCAS. These
datasets provide information on insurance contracts and claims related
to French motor third-party liability insurance. By leveraging Poisson
regression, our goal is to model the frequency of claims and investigate
the factors influencing claim occurrence within the insurance data.
:::

## Required Packages

```{r libraries, message=FALSE}

required_libraries <- c(
  "tidyverse", 
  "CASdatasets",
  "glmnet",
  "AER",
  "broom",
  "knitr",
  "kableExtra"
)
invisible(lapply(required_libraries, library, character.only = TRUE))
```


## Data

::: justify
The Data used in this vignette come from French motor third party
liability insurance portfolio.

The first dataset, `freMTPLfreq`, encompasses details regarding
contracts and clients obtained from a French insurance company, related
to some motor insurance portfolio.

The second dataset, `freMTPLsev`, contains claims information, from the
same company.

For convenience, the `freMTPLfreq` table will be named `CONTRACTS`
hereafter. The `freMTPLsev` will be named `CLAIMS`
:::

