---
title: "QuasiPoisson frequency analysis of a French MTPL dataset"
date: '`r format(Sys.Date(), "%B %Y")`'
author: "Julien Siharath"
chapters:
  - Introduction
  - Models
  - Graphics
categories: [Quasi-Poisson, Claim frequency, France, freMTPL]
bibliography: references.bib
fig-cap-location: top
editor: 
  markdown: 
    wrap: 72
format:
  html:     
    code-fold: true
    code-summary: "Show the code"
    css: style.scss
    theme: cosmo
    toc: true
    toc-depth: 2
---


# Introduction {#sec-quasi-poisson-introduction}

```{r setup, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Session Settings"
#| label: load-packages

# Graphs----
face_text='plain'
face_title='plain'
size_title = 14
size_text = 11
legend_size = 11

global_theme <- function() {
  theme_minimal() %+replace%
    theme(
      text = element_text(size = size_text, face = face_text),
      legend.position = "bottom",
      legend.direction = "horizontal", 
      legend.box = "vertical",
      legend.key = element_blank(),
      legend.text = element_text(size = legend_size),
      axis.text = element_text(size = size_text, face = face_text), 
      plot.title = element_text(
        size = size_title, 
        hjust = 0.5
      ),
      plot.subtitle = element_text(hjust = 0.5)
    )
}

# Outputs
options("digits" = 2)
```

```{css setup, echo = FALSE}
.justify {
  text-align: justify !important
}
```

::: callout-tip
### In Brief

The objective of this vignette is to demonstrate the application of Quasi-Poisson
regression in analyzing insurance data, specifically focusing on
the `freMTPLfreq` and `freMTPLsev` datasets from @charpentierCAS. These
datasets provide information on insurance contracts and claims related
to French motor third-party liability insurance. By leveraging Quasi-Poisson
regression, our goal is to model the frequency of claims and investigate
the factors influencing claim occurrence within the insurance data.
:::

## Required Packages

```{r libraries, message=FALSE}

required_libraries <- c(
  "tidyverse", 
  "CASdatasets",
  "wesanderson",
  "MASS",
  "AER",
  "broom",
  "knitr"
)
invisible(lapply(required_libraries, library, character.only = TRUE))
```

## Data

::: justify
The Data used in this vignette come from French motor third party
liability insurance portfolio.

The first dataset, `freMTPLfreq`, encompasses details regarding
contracts and clients obtained from a French insurance company, related
to some motor insurance portfolio.

The second dataset, `freMTPLsev`, contains claims information, from the
same company.

For convenience, the `freMTPLfreq` table will be named `CONTRACTS`
hereafter. The `freMTPLsev` will be named `CLAIMS`
:::

## Dictionaries

The list of the 10 variables from the `freMTPLfreq` dataset is reported
in @tbl-dict-freMTPLfreq. Similarly, the two variables encapsulated
within the freMTPLsev dataset are itemized in the table denoted as
@tbl-dict-freMTPLsev.

::: justify
| Attribute | Type    | Description                                             |
|----------------|----------------|---------------------------------------|
| PolicyID  | Numeric | Unique identifier for the contract                      |
| ClaimNB   | Numeric | Number of claims during the exposure period             |
| Exposure  | Numeric | Exposure duration in years                              |
| Power     | Factor  | Power of the car (ordered categorical)                  |
| CarAge    | Numeric | Age of the car in years                                 |
| DriverAge | Numeric | Age of the driver in years                              |
| Brand     | Factor  | Brand of the car categorized                            |
| Gaz       | Factor  | Type of fuel used (diesel or regular)                   |
| Region    | Factor  | Region in France where the driver resides               |
| Density   | Numeric | Population density in the city where the driver resides |

: Content of the `CONTRACTS` dataset: contracts {#tbl-dict-freMTPLfreq .striped .hover}

| Attribute   | Type    | Description                                 |
|-------------|---------|---------------------------------------------|
| PolicyID    | Numeric | Unique identifier for the contract          |
| ClaimAmount | Numeric | Cost of the claim, seen as at a recent date |

: Content of the `CLAIMS` dataset: claims {#tbl-dict-freMTPLsev .striped .hover}
:::

## Importation

```{r load-quasi-poisson-data, output=FALSE}
#| code-fold: true
#| code-summary: "code for importing our datasets"

data("freMTPLfreq")
data("freMTPLsev")

CONTRACTS <- freMTPLfreq |>
  filter(Exposure > 0.90)


# Create factors
CONTRACTS.f <- 
  CONTRACTS |> 
  mutate(
    DriverAge = cut(DriverAge, c(17, 22, 26, 42, 74, Inf)),
    CarAge = cut(CarAge, c(0, 1, 4, 15, Inf), include.lowest = TRUE),
    Density = cut(Density, c(0, 40, 200, 500, 4500, Inf), include.lowest = TRUE)
  )


CLAIMS <- freMTPLsev |>
  as_tibble()

# check
if (sum(freMTPLsev$PolicyID %in% CONTRACTS$PolicyID) == sum(CONTRACTS$ClaimNb)) {
  message("L'importation des données a été effectuée avec succès.")
} else {
  stop("Erreur lors de l'importation des données : les nombres ne sont pas égaux.")
}

```

# Models {#sec-quasi-poisson-models}

## Purpose

::: justify
In the realm of automobile insurance, Quasi-Poisson regression stands as a
reliable tool for understanding and predicting accident frequencies,
repair costs, and claims trends.

With Quasi-Poisson regression, insurers can anticipate forthcoming challenges,
refine pricing strategies, and ensure resilience in a dynamic landscape
of risk.
:::

::: {.callout-caution appearance="simple"}
## Pay Attention

::: justify
The results from QuasiPoisson regression models are valid if:\
- the responses are independent.\
- the responses are distributed according to a Poisson distribution with
parameter Lambda.\
- There may be [overdispersion](https://en.wikipedia.org/wiki/Overdispersion) present in the data.
Quasi-Poisson regression models are appropriate for handling situations where the variance exceeds the mean in the data.
:::
:::

The results from QuasiPoisson regression models are valid if:

The responses are independent.
The responses are distributed according to a Poisson distribution with parameter Lambda.

::: justify
In this analysis, we will explore the relationship between the response
variable `ClaimNb` and the explanatory variables `DriverAge` and
`Density`. This modeling framework aligns with the principles outlined
by @agresti, a prominent figure in statistical methodology, who
emphasizes the significance of considering multiple explanatory factors
in regression analysis.

The relationship under study can be represented by the following
equation:

$$
\begin{equation}
\text{ClaimNb} \sim \text{QuasiPoisson}(e^{X\beta + \log E})
\end{equation}
$$

$$
\begin{equation}
\log(\lambda) = \beta_0 + \beta_1 \times \text{DriverAge} + \beta_2 \times \text{Density} + \text{log(Exposure)}
\end{equation}
$$
:::

The estimated lambda parameter which is the mean of claims is:
`r mean(CONTRACTS.f$ClaimNb)`.

```{r lambda_parameter, lambda_parameter}

set.seed(1234) 

theoretic_count <- rpois(nrow(CONTRACTS.f), mean(CONTRACTS.f$ClaimNb))

tc_df <- tibble(theoretic_count)

freq_theoretic <- prop.table(table(tc_df$theoretic_count))

freq_claim <- prop.table(table(CONTRACTS.f$ClaimNb))

freq_theoretic_df <- tibble(

  Count = as.numeric(names(freq_theoretic)),
  Frequency = as.numeric(freq_theoretic),
  Source = "Theoretical Count"
)

freq_claim_df <- tibble(
  Count = as.numeric(names(freq_claim)),
  Frequency = as.numeric(freq_claim),
  Source = "Empirical Count"
)

freq_combined <- freq_theoretic_df |> 
  rbind(freq_claim_df)
```

The theoretical and empirical histograms associated with a Poisson
distribution are sown in @fig-plot-hist-claims.

```{r poisson-claim}
#| fig-cap: "<span style='color: #1E88E5;'>Theoretical</span> and <span style='color: black;'>empirical</span> histogram of claims in frequence"
#| label: "fig-plot-hist-claims"
#| code-fold: true
#| code-summary: "Code for the following graph"
#| fig-width: 8
#| fig-height: 5

ggplot(freq_combined, aes(x = Count, y = Frequency, fill = Source)) +
  geom_bar(stat = "identity", position = "dodge2", width = 0.3) +
  labs(x = "Claim Number", y = "Frequency", fill = "Legend") +
  theme(legend.position = "right") +
  scale_fill_manual(
    NULL,
    values = c("Empirical Count" = "black", "Theoretical Count" = "#1E88E5")
  ) +
  labs(fill = "Legend") +
  labs(x = "Claim Number", y = NULL) +
  theme(legend.position = "right")+
  global_theme()
```

```{r model}

freg <- formula(ClaimNb ~ DriverAge + Density + offset(log(Exposure)))
  
reg <- glm(freg, family = quasipoisson, data = CONTRACTS.f)

summary(reg)
```

::: justify
This is a Quasi-Poisson regression model predicting `ClaimNb` (number of
claims) with `DriverAge` and `Density` as predictors. The coefficients indicate the change in log count of claims
associated with each predictor level compared to a reference level. For
instance, as `DriverAge` increases from 22 to 26, the log count of
claims decreases by `r abs(round(coef(reg)[2], 2))` Similarly, as `Density` increases within
each category, the log count of claims increases.

All coefficients are statistically significant.
:::

::: panel-tabset
## Coefficients

```{r Coefficients, warning = FALSE}
#| tbl-cap: Coefficients
#| label: tbl-poisson-coef
#| code-fold: true
#| code-summary: code to create the table

reg_coef <- tidy(reg)

reg_coef$p.value <- format(reg_coef$p.value, scientific = TRUE, digits = 3)

kable(reg_coef)

```

## Count-Ratio

```{r count-ratio, warning = FALSE}
#| tbl-cap: Count Ratio
#| label: tbl-poisson-count-ratio
#| code-fold: true
#| code-summary: code to create the table

kable(exp(coef(reg)[-1]))

```

::: justify
Each count ratio represents the change in count of making a claim
associated with a one-unit increase in the predictor variable, compared
to the reference category `DriverAge` (18,22\]. For example, a count ratio of `r exp(coef(reg)[2])` for
`DriverAge` (22,26\] implies that the count of making a claim for
individuals aged 22 to 26 are approximately `r round((1-exp(coef(reg)[2]))*100)`% lower compared to the
reference category. Similarly, count ratios above 1 for `Density`
categories suggest an increase in the count of making a claim as the
density increases within each category.
:::

## Confidence intervals

```{r Conf, warning = FALSE}
#| tbl-cap: Confidence intervals
#| label: tbl-poisson-ci
#| code-fold: true
#| code-summary: code to create the table below

kable(exp(confint(reg))[-1, ])
  
```
:::

# Graphics {#sec-quasi-poisson-graphics}

::: panel-tabset
## DriverAge {#sec-quasi-poisson-graphics-age}

```{r Visualisation-coef-age, message = FALSE}
#| fig-cap: "Count ratio and confidence interval of DriverAge"
#| label: "fig-plot-coef-age"
#| code-fold: true
#| code-summary: code to create the following graph
#| fig-width: 8
#| fig-height: 5

count_ratio <- exp(coef(reg)[-1])
conf_int <- exp(confint(reg))[-1, ]

driver_age_vars <- grep("^DriverAge", names(count_ratio), value = TRUE)

data_age <- tibble(
  variable = driver_age_vars,
  coefficient = count_ratio[driver_age_vars], 
  lower_bound = conf_int[driver_age_vars, 1], 
  upper_bound = conf_int[driver_age_vars, 2]
)


driver_density_vars <- grep("^Density", names(count_ratio), value = TRUE)

data_density <- tibble(
  variable = driver_density_vars,
  coefficient = count_ratio[driver_density_vars], 
  lower_bound = conf_int[driver_density_vars, 1], 
  upper_bound = conf_int[driver_density_vars, 2]
)

ggplot(
  data_age, 
  aes(
    x = coefficient,
    y = rev(variable),
    xmin = lower_bound,
    xmax = upper_bound
  )
) +
  geom_point(stat = "identity", size = 3, color = "#1E88E5") +
  geom_errorbar(
    width = 0.2,
    position = position_dodge(width = 0.6),
    color = "#1E88E5"
  ) +
  labs(
    x = NULL,
    y = NULL
  ) +
  global_theme()
```

## Density {#sec-quasi-poisson-graphics-density}

```{r visualisation-coef-density}
#| fig-cap: "Count ratio and confidence interval of Density "
#| label: "fig-plot-coef-density"
#| code-fold: true
#| code-summary: code to create the following graph
#| fig-width: 8
#| fig-height: 5

data_density <- data_density |> 
  mutate(variable = reorder(variable, coefficient, decreasing = TRUE))

ggplot(
  data_density, 
  aes(
    x = coefficient,
    y = variable,
    xmin = lower_bound,
    xmax = upper_bound
  )
) +
  geom_point(
    stat = "identity",
    size = 3,
    color = "#1E88E5"
  ) +
  geom_errorbar(
    width = 0.2,
    position = position_dodge(width = 0.6),
    color = "#1E88E5"
  ) +
  labs(
    x = NULL,
    y = NULL
  ) +
  global_theme()
```
:::

# References

::: {#refs}
:::
