---
title: "Reserving exercice on reporting delays in AIDS Data"
date: '`r Sys.Date()`'
date-format: "MMMM, YYYY"
author: "Julien Siharath"
chapters:
  - Introduction
  - Overview
  - Models
  - Graphs
categories: [Triangle, Reserving, Reporting Delays, England, AIDS]
bibliography: references.bib
editor:
  markdown:
    wrap: 72
format:
  html:     
    embed-resources: true
    code-fold: true
    code-summary: "Show the code"
    css: style.scss
    theme: cosmo
    toc: true
    toc-depth: 2
vignette: >
  %\VignetteIndexEntry{Reserving exercice on reporting delays in AIDS Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

# Introduction {#sec-triangles-aids-introduction}

```{r setup, message=FALSE, warning=FALSE}
#| code-fold: true
#| code-summary: "Session Settings"
#| label: load-packages

# Graphs----
face_text='plain'
face_title='plain'
size_title = 14
size_text = 11
legend_size = 11

```

```{css setup, echo = FALSE}
.justify {
  text-align: justify !important
}
```

::: callout-tip
### In Brief

The reported number of Acquired Immune Deficiency Syndrome (AIDS) cases in England and Wales has recently been significantly underestimated due to substantial reporting delays. To address this issue, we will utilize the `Chainladder` package, specifically employing the `MackChainLadder` function. 

This analysis is crucial for accurately understanding the progression and impact of AIDS, which is essential for effective resource allocation and public health planning. By applying this method to AIDS reports in England and Wales from July 1983 to December 1992, we aim to adjust for reporting delays, thereby providing a more accurate picture of the epidemic's evolution during this period.

:::

## Required Packages

```{r libraries, message=FALSE}

required_libraries <- c(
  "tidyverse", 
  "tidyr",
  "ChainLadder",
  "boot",
  "dplyr"
)
invisible(lapply(required_libraries, library, character.only = TRUE))
```

## Data

::: justify

The AIDS dataset consists of 570 rows and 6 columns, capturing critical information about AIDS cases in England and Wales. Although all AIDS cases are required to be reported to the Communicable Disease Surveillance Centre, there is often a significant delay between diagnosis and reporting. To accurately estimate the prevalence of AIDS, it is essential to account for cases that have been diagnosed but not yet reported.

This dataset, obtained from @deangelis1994estimating and available in the `boot` package, records reported AIDS cases from July 1983 through December 1992. The data is organized by both the date of diagnosis and the delay in reporting, allowing for a more comprehensive analysis of the reporting delays and their impact on understanding the true prevalence of AIDS.

:::

## Dictionaries

The list of the 6 attributes from the `aids` dataset is reported
in @tbl-dict-aids.

::: justify
| Column Name | Description | Notes |
|-------------|-----------------------|---------------------------------------------|
| year        | The year in which the diagnosis was made.|                          |
| quarter     | The quarter of the year in which the diagnosis was made.| Values range from 1 to 4, representing Q1 to Q4.|
| delay       | The time delay (in months) between diagnosis and reporting.                                         | 0 indicates reporting within one month. Longer delays are grouped in 3-month intervals, with the value representing the midpoint of the interval (e.g., 2 means the report was delayed between 1 and 3 months). |
| dud         | An indicator of censoring, where categories have incomplete information.                            | The recorded number is a lower bound only.                                                         |
| time        | The number of quarters from July 1983 until the end of the quarter in which the cases were diagnosed. | |
| y           | The number of AIDS cases reported.       |    |


: Content of the `aids` {#tbl-dict-aids
.striped .hover}


:::


## Importation and triangle transformation

```{r load-triangles-data, output=FALSE}
#| code-fold: true
#| code-summary: "Code for importing the dataset and transforming it into a triangle"

data("aids")

aids_agg <- aids |>
  group_by(year, quarter, delay) |>
  summarise(cases = sum(y), .groups = 'drop') |>
  mutate(year_quarter = paste(year, quarter, sep = "-"))

# Create a data frame in long format
aids_long <- aids_agg |>
  select(year_quarter, delay, cases)

# Reshape the data into a wide format (triangle format)
triangle <- aids_long |>
  pivot_wider(names_from = delay,
              values_from = cases,
              values_fill = list(cases = 0)) |>
  arrange(year_quarter)

# Convert the wide format data frame to a matrix
  triangle_matrix <- as.matrix(triangle |>
  select(-year_quarter))  # Exclude the year_quarter column for matrix conversion

rownames(triangle_matrix) <- triangle$year_quarter

# Triangular shape
n_rows <- nrow(triangle_matrix)
n_cols <- ncol(triangle_matrix)
  
for (i in seq_len(n_cols-1)) {
    triangle_matrix[(n_rows - i + 1):n_rows, i+1] <- NA
}

full_rows <- which(rowSums(is.na(triangle_matrix)) == 0)
full_triangle <- triangle_matrix[full_rows, ]
close_data <- colSums(full_triangle, na.rm = TRUE)

triangle_rows <- which(rowSums(is.na(triangle_matrix)) != 0)

triangle_matrix_final <- rbind(close_data, triangle_matrix[triangle_rows, ])

# Convert matrix to a 'triangle' object for chainladder
triangle_cl <- as.triangle(triangle_matrix_final)

print(triangle_cl)

```
In this context, a triangle is a table used to display data over time, organized across two dimensions:

- **Lines as Origin Year:** Represents the year when the diagnosis occurred.
- **Columns as Development Year:** Indicates the number of months (+/- 1) that have passed since the origin year.

This table structure is instrumental in tracking and analyzing the progression of data from the time of origin over subsequent months. It allows for a clear visualization of how cases develop and accumulate over time, providing valuable insights for further analysis.

# Overview {#sec-triangles-aids-overview}

## Purpose

::: justify

The prediction and monitoring of the acquired immune deficiency syndrome (AIDS) epidemic rely heavily on the availability of reliable and complete data on AIDS diagnoses. However, a significant challenge in this context is the considerable delay that often occurs between the diagnosis of an AIDS case and its subsequent reporting to the epidemic monitoring center. In England and Wales, this reporting process is managed by the Public Health Laboratory Service AIDS Centre at the Communicable Disease Surveillance Centre. To effectively utilize AIDS reports, it is crucial to account for cases that have been diagnosed but have not yet been reported.

To address this issue, we will employ the Mack chain ladder method, a statistical tool originally developed for insurance claims forecasting. The method is particularly well-suited for analyzing right-truncated data, making it an ideal choice for correcting delays in reporting. The Mack chain ladder method enables robust predictions of future case reports by estimating the development pattern of reported cases over time. This approach not only adjusts for the reporting delay but also provides a measure of the uncertainty associated with these estimates.

By applying the Mack chain ladder method, we aim to achieve more accurate and timely insights into the progression of the AIDS epidemic. Enhanced prediction and monitoring capabilities are essential for effective public health planning, resource allocation, and the implementation of timely interventions to control the spread of the disease. While various statistical techniques have been developed to address this inferential challenge, the Mack chain ladder method is distinguished by its reliability and practicality in handling right-truncated data.
:::

::: panel-tabset

## Claims development chart
```{r dev-chart}
#| code-fold: true

triangle_cum <- incr2cum(triangle_cl)
print(triangle_cum)
```

```{r TRIANGLES-plot}
#| fig-cap: "Claims development chart of the damage triangle, with one line per origin period."
#| label: "fig-dev-chart"
#| code-fold: true
#| code-summary: Code to create the following graph
#| fig-width: 8
#| fig-height: 5

ChainLadder::plot(as.triangle(triangle_cum[-1, -n_cols]))
```
## Origin year plot

```{r}
#| fig-cap: "Claims development by origin year"
#| label: "fig-dev-origin"
#| code-fold: true
#| code-summary: Code to create the following graph
#| fig-width: 8
#| fig-height: 5

plot(as.triangle(triangle_cum[-1, -n_cols]), lattice=TRUE)
```
:::

# Mack chain-ladder

::: justify
::: justify
In 1993, Thomas Mack introduced a groundbreaking method in his paper [@Mack_distributionfree1993], which estimates the standard errors of the chain-ladder forecast without assuming a specific distribution, under three key conditions.

## The Mack Chain-Ladder Model

Following the notation established by Mack in 1999 [@Mack1999], let $C_{ik}$ denote the cumulative loss amounts of origin period (e.g., accident year) $i=1,\ldots,m$, with losses known for development period (e.g., development year) $k \le n+1-i.$

To forecast the amounts $C_{ik}$ for $k > n+1-i$, the Mack chain-ladder model makes the following assumptions:

$$
\begin{aligned}
  \text{CL1:} & \quad E[ F_{ik}| C_{i1},C_{i2},\ldots,C_{ik} ] = f_k
  \quad \text{where} \quad F_{ik} = \frac{C_{i,k+1}}{C_{ik}} \\
  \text{CL2:} & \quad Var\left( \frac{C_{i,k+1}}{C_{ik}} \Bigg| C_{i1},C_{i2}, \ldots,C_{ik} \right) = \frac{\sigma_k^2}{w_{ik} C_{ik}^\alpha} \\
  \text{CL3:} & \quad \{C_{i1},\ldots,C_{in}\} \text{ and } \{C_{j1},\ldots,C_{jn}\} \text{ are independent for origin periods } i \neq j
\end{aligned}
$$

where $w_{ik} \in [0,1]$ and $\alpha \in \{0,1,2\}$ are parameters that adjust the variance structure. If these assumptions hold, the Mack chain-ladder model provides an unbiased estimator for Incurred But Not Reported (IBNR) claims.

## Intuition Behind the Method

The Chain-Ladder model is a powerful tool used to project future claims based on historical data. The core idea is that claims development patterns are relatively stable and predictable. The model assumes that the ratios of cumulative losses between successive development years are consistent, allowing these ratios to be used in estimating future losses.

### Assumptions Explained:

1. **CL1: Expected Future Ratio**  
   This assumption posits that the expected future ratio $F_{ik}$ of cumulative losses between successive development years is constant, given the known losses up to the current development year. Essentially, this means that the ratio of cumulative losses from one development year to the next is assumed to follow a consistent pattern, captured by a factor $f_k$.

2. **CL2: Variance of Future Ratio**  
   Here, the model specifies that the variance of the future ratio of cumulative losses is proportional to $\frac{\sigma_k^2}{w_{ik} C_{ik}^\alpha}$. The term $\sigma_k^2$ represents the variability, $w_{ik}$ is a weight, and $\alpha$ is a parameter that adjusts for different variance levels. This assumption is crucial for quantifying the uncertainty around the forecasts.

3. **CL3: Independence of Origin Periods**  
   This assumption ensures that cumulative loss amounts from different origin periods (e.g., different accident years) are independent. This independence simplifies the estimation process and increases the robustness of the model.

### Practical Application

The Mack Chain-Ladder model can be viewed as a weighted linear regression through the origin for each development period:

$$
\text{lm}(y \sim x + 0, \text{weights} = w/x^{2-\alpha}),
$$

where $y$ is the vector of claims at development period $k+1$ and $x$ is the vector of claims at development period $k$.

The Mack method is implemented in the ChainLadder package via the function `MackChainLadder`. This implementation enables actuaries to perform robust reserving exercises, forecast future claims developments, and maintain the financial stability of insurance companies by ensuring they can meet their future claim obligations.

For a comprehensive understanding of this methodology, including its practical implications and applications, see @gesmann2014claims.

As an example we apply the `MackChainLadder` function to our
triangle `Damage`:
::: 

```{r MackChainLadder}
#| code-fold: false

mack <- MackChainLadder(triangle_cum, est.sigma="Mack")
mack # same as summary(mack) 
```
```{r}
#| code-fold: false
#| code-summary: Show the code

# Displaying the Mack model's parameters
mack$f

# Viewing the full triangle data from the Mack model
mack$FullTriangle
```
:::

::: justify

The Mack model factors start at 3.8 for the initial periods and gradually decrease to around 1 in the later periods. These factors represent the development factors for each period, indicating the expected growth in cumulative reports from one period to the next. This trend reflects the typical pattern where most reports occur early in the development process, with the rate of increase diminishing over time.

The triangular data illustrates how reports develop over time for each origin period. For instance, in the 4th quarter of 1992, the number of reported cases of acquired immune deficiency syndrome (AIDS) begins at 67 in the 1st period and grows to 437 cases by the 41st period.

Overall, this output offers a comprehensive view of how reports develop over time and highlights the associated development factors used in the chain ladder method.

:::
# References

::: {#refs}
:::

# See also

::: justify
For more similar triangles datasets, see
[`nortritpl8800`](https://dutangc.github.io/CASdatasets/reference/nortritpl.html)
(import with `data("ausprivauto0405")`): Australian liabilty insurance triangles dataset,
[`sgautoprop9701`](https://dutangc.github.io/CASdatasets/reference/sgtriangles.html):
Singapore general liability triangles dataset (import with `data("norauto")`),
[`swtri1auto`](https://dutangc.github.io/CASdatasets/reference/swtriangles.html):
Switzerland general liability triangles dataset (import with `data("beMTPL16")`), or
[`usautotri9504`](https://dutangc.github.io/CASdatasets/reference/usautotri.html)
(import with `data("pg17trainpol")`): US Automobile triangles dataset.
:::
